Accessible and Completed Datasets yet Majority are Bronze-Graded and not Updated Regularly with Missing Metadata

### 2025-05-16 paper/paper.qmd
Made fixes

In this paper, we analyzed data provided by Open Data Toronto on the data quality grade of the datasets in their catalogue and their characteristics like "accessibility", "completeness", "freshness", "metadata", and "usability". Farrow analyzed the data quality scores of the different datasets on the Open Data Toronto portal and found that the metadata and freshness scores were poor overall [@citeFarrow]. However there has been no analysis done on the characteristics and how they compare for different graded datasets as of 2025. In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimial missing data and are usable. It also showed that 56% of their datasets are graded bronze with bronze-graded datasets being less likely to be updated and have completed metadata fields. These bronze-graded datasets contribute the poor metadata and freshness scores we saw. These findings can inform Open Data Toronto about specific graded datasets that should be given more attention and the specific qualities of them such as metadata completion, which impacts the quality of the data used in the media and in policymaking. These findings can also be informative for users of the portal in order to understand what goes behind the grade of datasets on the portal and what they mean.

@@

Our variables of interest that we used in our analysis are the following: "accessibility", "completeness", "freshness", "metadata", "usability", and "grade". "Accessiblity" is a score from 0 to 1 that indicates the degree that the dataset can be access or not through the Open Data Toronto API, keywords or tags, and automated data pipelines with 1 being that it can not be accessed with the various methods noted and 0 if not at all. "Completeness" is a score from 0 to 1 that indicates how much of the data is missing with 1 being that there is no missing data and 0 being that all data fields are empty or missing. The "freshness" variable is a score from 0 to 1 that indicates how up-to-date the data where a shorter time duration between the recent refresh date and time and the previous refresh date and time before that gives a higher score. The "metadata" variable indicates how complete the following metadata fields are for a dataset from a scale of 0 to 1: Description, Limitations, Topics, Contact Email. The more metadata fields compelted, the higher the metadata score. The "usability" variable is a score from 0 to 1 indicating how easy it would be to use the dataset and this is determine by the proportion of meaningful column names or in other words column names with English words in it. However a limitation of the "usability" variables is that it does not consider datasets that are in different languages other than English.

>>>>

In this paper, we analyzed data provided by Open Data Toronto on the data quality grade of the datasets in their catalogue and their characteristics like "accessibility", "completeness", "freshness", "metadata", and "usability". Farrow analyzed the data quality scores of the different datasets on the Open Data Toronto portal and found that the metadata and freshness scores were poor overall [@citeFarrow]. However there has been no analysis done on the characteristics and how they compare for different graded datasets as of 2025. In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimal missing data and are usable. It also showed that 56% of their datasets are graded bronze with bronze-graded datasets being less likely to be updated and have completed metadata fields. These bronze-graded datasets contribute the poor metadata and freshness scores we saw. These findings can inform Open Data Toronto about specific graded datasets that should be given more attention and the specific qualities of them such as metadata completion, which impacts the quality of the data used in the media and in policymaking. These findings can also be informative for users of the portal in order to understand what goes behind the grade of datasets on the portal and what they mean.

@@

Our variables of interest that we used in our analysis are the following: "accessibility", "completeness", "freshness", "metadata", "usability", and "grade". "Accessibility" is a score from 0 to 1 that indicates the degree that the dataset can be access or not through the Open Data Toronto API, keywords or tags, and automated data pipelines with 1 being that it can not be accessed with the various methods noted and 0 if not at all. "Completeness" is a score from 0 to 1 that indicates how much of the data is missing with 1 being that there is no missing data and 0 being that all data fields are empty or missing. The "freshness" variable is a score from 0 to 1 that indicates how up-to-date the data where a shorter time duration between the recent refresh date and time and the previous refresh date and time before that gives a higher score. The "metadata" variable indicates how complete the following metadata fields are for a dataset from a scale of 0 to 1: Description, Limitations, Topics, Contact Email. The more metadata fields completed, the higher the metadata score. The "usability" variable is a score from 0 to 1 indicating how easy it would be to use the dataset and this is determine by the proportion of meaningful column names or in other words column names with English words in it. However a limitation of the "usability" variables is that it does not consider datasets that are in different languages other than English.
